---
title: "pamlApps demo code"
author: "Janet Young\n"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(patchwork)
library(kableExtra)

library(ape) # for plotTree function to plot branch paml results

library(treeio) # for parseMLCbranches_new function to read branch paml results
library(ggtree) # for plotTree_new function to plot branch paml results

source("parseAndPlotPAML_functions.R")
```

# sitewise PAML

Read a model 8 rst file:
```{r}
ACE2_sitePAML_rstFile <- "data/paml_version_4.10.6/example_ACE2/M8_initOmega0.4_codonModel2/rst"

## get the BEB table as a data.frame
ACE2siteResults <- parseRSTfile(ACE2_sitePAML_rstFile) # %>% 
#  as_tibble()
```

look at that table for sites where BEB >= 0.90

```{r}
ACE2siteResults[which(ACE2siteResults[,"prob_class11_omega_4.69678"] >= 0.90),] %>% 
    as_tibble() %>% 
    dplyr::select(pos, human_ACE2_ORF, meanOmega, prob_class11_omega_4.69678)
```


plot BEB mean dN/dS estimate

```{r, fig.height=4, fig.width=9}
plotOmegas(ACE2siteResults, 
           title="ACE2 model 8 BEB mean dN/dS estimates", 
           highlightHighBEB=TRUE, 
           highBEBthreshold=0.90, 
           highBEBcolor="red")
```


plot BEB mean dN/dS estimate in a visually unusual way, where bars all start at dN/dS=1. Our tetherin collaborators liked seeing the results this way but it is not intuitive: makes it look like most sites have dN/dS=1.

```{r, fig.height=4, fig.width=9}
plotOmegas(ACE2siteResults, 
           title="ACE2 primates: model 8 BEB mean dN/dS estimates", 
           highlightHighBEB=TRUE, 
           highBEBthreshold=0.90, 
           highBEBcolor="red", 
           yAxisCenterAtNeutral=TRUE)
```


plot BEB probabilities of being in the positively selected class

```{r, fig.height=4, fig.width=9}
plotProbs(ACE2siteResults, 
          title="ACE2 primates: M8 BEB probabilities of positive selection",
          addThresholdLine=TRUE, 
          threshold=0.90, 
          thresholdLineColor="red",
          highlightHighBEB=TRUE, 
          highBEBthreshold=0.90, 
          highBEBcolor="blue")
```

# Branch PAML

## using original functions (base R)

Read branch PAML's main output file, which will be called `mlc` unless you specified otherwise.  The `parseMLCbranches()` function returns a list with two objects, named `tree` (a phylo object) and `table` (a data.frame containing all results reported in the tabular section of PAML's mlc output file). 

```{r}
ACE2_branchPAML_mlcFile <- "data/paml_version_4.10.6/example_ACE2/BRANCHpaml_initOmega0.4_codonModel2/mlc"

ACE2_branchPAML_results <- parseMLCbranches(ACE2_branchPAML_mlcFile)
```

Use the output of `parseMLCbranches()` to plot the tree, showing N, S, and dN/dS along each branch.

The default is to color the label for any branch where dN/dS > 1.

```{r, fig.height=9, fig.width=9}
plotTree(phymlTree=ACE2_branchPAML_results[["tree"]], 
         mlcTable=ACE2_branchPAML_results[["table"]], 
         labelType="omega_NS",
         myTitle="ACE2 branch PAML")
```


I can change the dN/dS threshold, as well as requiring a minimum N*dN (number of non-synonymous changes), and I can use a different color:


```{r, fig.height=9, fig.width=9}
plotTree(phymlTree=ACE2_branchPAML_results[["tree"]], 
         mlcTable=ACE2_branchPAML_results[["table"]], 
         labelType="omega_NS",
         myTitle="ACE2 branch PAML", 
         colorHighOmega=TRUE, 
         colorHighOmegaThreshold=1.5,
         colorHighOmegaThresholdNxN=3,
         highOmegaColor="orange")
```


## using newer functions (uses ggtree), allows rerooting


```{r, fig.height=9, fig.width=15}
ACE2_branchPAML_results_treeio_v2 <- parseMLCbranches_new(ACE2_branchPAML_mlcFile)

ACE2_branchPAML_results_treeio_v2_reroot <- root(
    ACE2_branchPAML_results_treeio_v2,
    node=getMRCA(ACE2_branchPAML_results_treeio_v2@phylo,
                 c("human_ACE2_ORF",
                   "white-tufted-ear_marmoset_ACE2")))

p1 <- plotTree_new(ACE2_branchPAML_results_treeio_v2, 
                   myTitle="ACE2 branch PAML, ggtree-based plot function")
p2 <- plotTree_new(ACE2_branchPAML_results_treeio_v2_reroot, 
                   myTitle="ACE2 branch PAML, rerooted, ggtree-based plot function")

p1 + p2
```


# Finished

show R version used, and package versions

```{r sessionInfo}
sessionInfo()
```


